---
title: "Interactive supplement to manuscript 'Exceptional trends in adolescent alcohol use in the Netherlands'"
author: "Rianne Margaretha Schouten, and others"
date: "18/08/2021"
output:
  html_document:
    css: custom.css
runtime: shiny
resource_files:
#- maxmovprevslope.PNG
#- maxmovprevslopecount.PNG
- maxprev.PNG
#- maxprevslope.PNG
---

<style type="text/css">

body{ /* Normal  */
      font-size: 18px;
  }
h2 { /* Header 2 */
    font-size: 26px;
  color: DarkBlue;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Introduction

## Methods

```{r, include=FALSE, results='hide', cache=TRUE}
library(reticulate)
library(tidyverse)
library(ggplot2)
library(Cairo)
library(randomcoloR)
library(shiny)
#options(shiny.usecairo=T)
```

```{r, include=FALSE, results='hide', cache=TRUE}
select_data <- function(name){

  if(name == 'maxprev'){
    data <- read.csv('./data/maxprev.csv', sep = ",")
  #} else if (name == 'maxprevslope'){
  #  data <- read.csv('./data/maxprevslope.csv', sep = ",")
  } else if (name == 'maxmovprevslope'){
    data <- read.csv('./data/maxmovprevslope.csv', sep = ",")
  } else if (name == 'maxmovprevslopecount'){
    data <- read.csv('./data/maxmovprevslopecount.csv', sep = ",")
  }
  
  data$subgroup <- as.factor(as.numeric(data$subgroup))
  data$year = rep(c(1:9), 20 + 1)
  return(data)
}
```

```{r, include=FALSE, results='hide', cache=TRUE}
palette <- c("#a6cee3", "#1f78b4", "#b2df8a", "#33a02c", "#fb9a99", "#e31a1c", "#fdbf6f", "#ff7f00",
             "#cab2d6", "#6a3d9a", "#ffff99", "#b15928", "#8dd3c7", "#fb8072", "#bc80bd", "#fccde5", 
             "#fdb462", "#d9d9d9", "#ccebc5", "#b3de69")
palette <- c(palette, "#636363")
make_plot <- function(data, sel_numbers, outcome){
  
  sel <- data[data$subgroup %in% c(sel_numbers,50), ] 
  colors <- palette[c(sel_numbers,21)]
  
  if(outcome == 'prev'){
    y = 'prev'
    breaks = c(1:9) #seq(2003, 2019, 2)
    x_labels = c('03', '05', '07', '09', '11', '13', '15', '17', '19')
  } else if(outcome == 'mov_prev'){
    y = 'mov_prev'
    x_labels = c('03/05', '05/07', '07/09', '09/11', '11/13', '13/15', '15/17', '17/19')
    breaks = 1:8
    sel <- sel %>% filter(year != 9)
  }
  
  trend_plot <- ggplot(sel, aes_string(x = 'year', y = y, color = 'subgroup')) + 
  geom_point(size=1.2) + 
  geom_line(size=1) + 
  ggtitle(label = "") + 
  scale_x_continuous(breaks=breaks, labels=x_labels) + 
  scale_y_continuous(breaks=seq(0.0,0.8,0.2), limits=c(0.0,0.85)) +
  xlab("") + 
  ylab("") + 
  scale_color_manual(values = colors, 
                     labels = c(sel_numbers, "D"),
                     name = "") + 
  guides(color = guide_legend(nrow=2, ncol=10, byrow=TRUE, 
                              override.aes = list(size = 0.4)),
         shape = guide_legend(override.aes = list(size = 0.4))) + 
  theme_bw(base_size=16) + 
  theme(legend.position="top",
        legend.justification="right",
        #legend.direction = "vertical",
        #plot.title = element_text(hjust = 0, vjust=-3), 
        #legend.box.margin = margin(0,0,-0.2,0, "line"),
        #axis.title.y = element_text(),
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        #panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        #legend.text  = element_text(size = 18),
        #legend.key.width = unit(1.2,"cm"),
        #legend.key.size = unit(1.2,"cm"),
        plot.margin = unit(x = c(-2, 1, -2, -2), units = "mm"))
  
  return(trend_plot)
}

make_table <- function(name,data){
  
  if(name == 'maxprev'){
    table <- data.frame(data)
    #table <- data.frame(row.names = c(1:20))
    #table$sg <- c('1','2','3','4','5','6','7','8','9', '10', '11', '12', '13', '14', '15', '16', '17', '18', '19', '20')
    #table$cov <- c('0.3','0.4', '0.4', '0.2','0.1','0.2','0.1','0.1', '0.1', '0.2', '0.3','0.4', '0.4', '0.2','0.1','0.2','0.1','0.1', '0.1', '0.1')
    #table$description <- c('age < 0.5 AND life satisfaction > 7and life satisfaction > 7',
  #)
  }
  
  return(table)
  
}
```

## Results {#results .emphasized}

Interactive results are presented in the plot below. 

- **Select subgroups:** fill in numbers between 1 and 20, use comma to separate

- **Type of exceptionality:** select one of three types as mentioned in manuscript

- **Y-axis variable:** select prevalence as trend variable

<p>&nbsp;</p>

```{r include=FALSE, results='hide', cache=TRUE}
extract <- function(text) {
    text <- gsub(" ", "", text)
    split <- strsplit(text, ",", fixed = FALSE)[[1]]
    as.numeric(split)
}
```

```{r }
plot <- renderPlot({
  data <- select_data(name = input$select_file)
  nums <- extract(input$subgroups)
  make_plot(data=data, sel_numbers=nums, outcome=input$yaxis)
}, res = 96, height=500, width=600)

table <- renderImage({
  if(input$select_file == "maxprev"){            
      list(src = "./www/maxprev.png")
  } else if(input$select_file == "maxmovprevslope"){
      list(src = "./www/maxmovprevslope.png")
  } else if(input$select_file == "maxmovprevslopecount"){
      list(src = "./www/maxmovprevslopecount.png")
  }}, deleteFile = FALSE)

#table <- div(renderTable(make_table(name = input$select_file,data=select_data(name = input$select_file))), style="font-size:30%")

fluidPage(
  fluidRow(
    
    #column(6, offset=0, style='padding:0px;margin:0px;',

       #fluidRow(
           
    column(3, offset=0, style='padding:0px;margin:0px;',
           textInput("subgroups", "Subgroups:", 
            placeholder = "Enter numbers separated by a comma...",
            value = "1,2,3")),
    
    column(3, offset=0, style='padding:0px;margin:0px;',
            selectInput("select_file", "Exceptionality:",
                c("1. Prevalence" = "maxprev",
                  #"Slope" = "maxprevslope",
                  "2. Average slope" = "maxmovprevslope",
                  "3. Horizontal average slope" = "maxmovprevslopecount"
                  ))),
    
    column(3, offset=0, style='padding:0px;margin:0px;',
            selectInput("yaxis", "Y-axis variable:",
                c("Prevalence" = "prev",
                  "Average of prevalence" = "mov_prev"
                  ),
                selected = 1))
  ),
  fluidRow(
         
    column(9, plot, offset=4, style='padding:0px;margin:0px;')
  ),
  #fluidRow(
  #  column(9, style = 'padding:0px;margin:0px;')
  #),
  #fluidRow(
  #  column(9, style = 'padding:0px;margin:0px;')
  #),
    #),
    #column(1, offset=0, style='padding:0px;margin:0px;'),
    #column(5, table, offset=0, style='padding:0px;margin:0px;')
    #column(6, offset=0, style='padding:0px;margin:0px;',
  fluidRow(
     column(9, table, offset=2, style='padding:0px;margin:0px;')
  )
)
```

---

